---
output: 
  github_document:
    # Controls whether an HTML preview file (README.html) is created alongside README.md
    # true: creates both .md and .html files
    # false: creates only .md file (recommended for GitHub repos)
    html_preview: false

    # Controls automatic table of contents generation
    # true: generates a table of contents based on headers
    # false: no automatic table of contents (you can still add one manually)
    toc: false
    
    # Controls how line breaks are handled
    # true: every line break in the .Rmd becomes a hard break in the .md
    # false: follows standard markdown where single line breaks are ignored 
    #        (need two line breaks for a paragraph break - this matches GitHub's behavior)
    hard_line_breaks: false

    # Controls how math expressions are handled
    # null: keeps LaTeX math as-is (recommended for GitHub)
    # "katex": renders using KaTeX
    # "webtex": renders using webtex
    math_method: null

    # Optional additional settings you might want to consider:
    
    # Controls the depth of headers included in TOC (if toc: true)
    # toc_depth: 3
    
    # Adds section numbers to headers (if toc: true)
    # number_sections: false
    
    # Controls whether code chunks are highlighted
    # highlight: true
    
    # Sets the theme for syntax highlighting
    # highlight_style: "github"
    
    # Controls whether figures are centered
    # fig_align: "center"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  warning = FALSE,
  message = FALSE
)

# Function to get vignette information and create links
get_vignette_links <- function() {
  # List all .Rmd files in vignettes directory
  vig_files <- list.files(path = "vignettes", 
                         pattern = "\\.Rmd$", 
                         full.names = TRUE)
  
  if (length(vig_files) == 0) return(NULL)
  
  # Extract titles and filenames
  vig_info <- lapply(vig_files, function(file) {
    # Read the YAML header
    lines <- readLines(file, n = 20)
    yaml_start <- which(lines == "---")[1]
    yaml_end <- which(lines == "---")[2]
    yaml_content <- lines[(yaml_start+1):(yaml_end-1)]
    
    # Extract title
    title_line <- grep("^title:", yaml_content, value = TRUE)
    title <- gsub("^title:\\s*|\"", "", title_line)
    
    # Get filename without path and extension
    basename <- tools::file_path_sans_ext(basename(file))
    
    list(file = basename, title = title)
  })
  
  # Convert to data frame
  vig_info <- do.call(rbind, lapply(vig_info, data.frame))
  
  # Sort by filename
  vig_info <- vig_info[order(vig_info$file), ]
  
  # Create markdown links
  base_url <- "https://benjaminpeeters.github.io/WPD/articles/"
  vig_links <- sprintf(
    "%d. [%s](%s%s.html)",
    seq_len(nrow(vig_info)),
    vig_info$title,
    base_url,
    vig_info$file
  )
  
  paste(vig_links, collapse = "\n")
}
```


# WPD: World Panel Data

[![R-CMD-check](https://github.com/benjaminpeeters/WPD/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/benjaminpeeters/WPD/actions)
[![Website](https://img.shields.io/badge/website-WPD-blue)](https://benjaminpeeters.github.io/WPD/)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html#maturing)
[![License: AGPL v3](https://img.shields.io/badge/License-AGPL%20v3-blue.svg)](https://www.gnu.org/licenses/agpl-3.0)

## Overview

WPD (World Panel Data) is an R package that simplifies access to and visualization of global economic and social panel data from major international organizations and academic researchers. 
It aims to serve as a bridge between publicly available datasets and researchers, students, and the general public, making cross-country panel data easier to work with.

The package includes a large database (approximately 50MB) of world panel data that is loaded on demand through the 'wp_data' function. 
Features include customizable time series plots, cross-sectional analysis tools, scatter plots with trend analysis, bar charts for comparative analysis, built-in seasonal adjustment capabilities, and support for both annual and quarterly frequencies. 
The package focuses on macroeconomic indicators such as GDP, trade, prices, employment, and financial and balance-of-payments data across multiple countries and regions.


<!-- Place overview.png here: A visual showing the package's workflow from data sources to visualization -->

### Key Features

- **Unified Data Access**: Load and manipulate data from multiple sources through a single interface
- **Data Harmonization**: Standardized country names and ISO3 codes across all sources
- **Advanced Data Processing**: 
  - Frequency matching (quarterly to yearly or vice versa)
  - Multiple interpolation methods
  - Seasonal adjustments for quarterly data
  - Group aggregation with various methods
  - Automatic data validation and cleaning
- **Publication-Ready Visualizations**:
  - Time series plots with multiple customization options
  - Scatter plots with statistical annotations
  - Bar plots with flexible grouping
  - Support for multi-panel layouts
  - Export in various formats (PNG, PDF)


<!-- Dynamic content: Documentation links -->
## Getting Started

Visit our [Documentation Website](https://benjaminpeeters.github.io/WPD/index.html) for comprehensive guides and tutorials. We recommend starting with:

```{r vignettes, echo=FALSE, results='asis', eval=TRUE}
# This chunk dynamically generates vignette links
cat(get_vignette_links())
```

Click here for [complete function reference](https://benjaminpeeters.github.io/WPD/reference/index.html).


### Installation

```r
# Install from GitHub
devtools::install_github("benjaminpeeters/WPD")
```

### Technical Requirements

- R version ≥ 3.5.0
- Disk space: ~100MB for installation
- Runtime memory: 15-50MB (depending on data loaded)
- Dependencies: ggplot2 (≥3.4.0, <3.5.0), scales, patchwork, zoo, seasonal

### Basic Example

```r
library(WPD)

# Load data for G7 countries
data <- wp_data(
  ISO = "G7",
  formula = c("GDP_C", "CU_C/GDP_C*100"),
  variable = c("GDP", "Current Account"),
  years = c(2000, 2023)
)

# Create a time series plot
wp_plot_series(data, 
               y_axis = c("Billion USD", "% of GDP"),
               right_axis = "Current Account",
               title = "GDP and Current Account in G7 Countries")
```

<!-- Place example1.png here: Output of the above code showing GDP and Current Account for G7 -->

## Available Data

Browse available data using the `wp_info()` function:

```r
# View all available variables
wp_info()

# View only quarterly data
wp_info("Q")

# View only yearly data
wp_info("Y")
```

<!-- Place data_coverage.png here: Visual showing temporal and geographical coverage of different data sources -->

## Advanced Features

### Data Processing

```r
# Complex data manipulation example
data <- wp_data(
  ISO = c("EU", "BRICS"),  # Multiple country groups
  formula = "100*CU_C/GDP_C",  # Custom formula
  years = c(2000, 2023),
  adjust_seasonal = TRUE,  # Apply seasonal adjustment
  aggregate_iso = "Mean",  # Average across countries
  aggregate_period = "CAGR"  # Calculate compound growth
)
```

### Visualization Examples

#### Time Series Plots

```r
# Multi-panel time series with events
events <- data.frame(
  Event = c("Global Financial Crisis", "Covid-19 Pandemic"),
  Date = as.Date(c("2008-09-15", "2020-03-11"))
)

wp_plot_series(data,
               key_dates = events,
               area = TRUE,  # Create area plot
               reference = TRUE)  # Add data sources
```

<!-- Place timeseries.png here: Example of multi-panel time series with event markers -->

#### Scatter Plots

```r
# Scatter plot with regression and grouping
wp_plot_scatter(data,
                color = "Region",
                interpolation = "Spline",
                r_squared = 3,  # Show full equation
                ISO = "Both")  # Show both points and labels
```

<!-- Place scatter.png here: Example scatter plot with regression line and country labels -->




<!-- Dynamic content: Citation -->
## Citation and Data Attribution

When using WPD, please cite both the package and the original data sources:

```{r citation, echo=TRUE, eval=TRUE, comment=""}
# Generate citation dynamically
citation("WPD")
```



## Data Sources and References

### International Organizations

1. **World Bank Development Indicators**  
   Comprehensive collection of development indicators.  
   https://data.worldbank.org/indicator

2. **International Monetary Fund**
   - International Financial Statistics (IFS)  
     Global financial and economic indicators.
   - Balance of Payments Statistics  
     International transactions data.

3. **Bank for International Settlements (BIS)**  
   Credit aggregates, property prices, and exchange rates.

4. **OECD Main Economic Indicators**  
   Economic indicators for OECD members and partners.

### Research Databases

5. **Jordà-Schularick-Taylor Macrohistory Database**
   - Jordà, Ò., Schularick, M., & Taylor, A. M. (2017). "Macrofinancial History and the New Business Cycle Facts." *NBER Macroeconomics Annual*, 31(1), 213-263.
   - Jordà, Ò., et al. (2019). "The Rate of Return on Everything, 1870–2015." *The Quarterly Journal of Economics*, 134(3), 1225-1298.
   - Jordà, Ò., et al. (2021). "Bank Capital Redux: Solvency, Liquidity, and Crisis." *Review of Financial Studies*, 34(7), 3062-3107.

6. **KOF Globalisation Index**  
   Gygli, S., Haelg, F., Potrafke, N., & Sturm, J. E. (2019). "The KOF Globalisation Index – Revisited." *Review of International Organizations*, 14(3), 543-574.

7. **Emergency Events Database (EM-DAT)**  
   Centre for Research on the Epidemiology of Disasters (CRED). (2024). EM-DAT: The International Disaster Database.

8. **Sovereign Default Database**  
   Beers, D., Ndukwe, C. I., & Charron, S. (2024). BoC–BoE Sovereign Default Database. Bank of Canada, Bank of England.

9. **Capital Account Openness**  
   Quinn, D., & Toyoda, A. M. (2008). "Does Capital Account Liberalization Lead to Economic Growth?" *Review of Financial Studies*, 21(3), 1403-1449.

10. **Maddison Project Database**  
    Maddison Project Database 2023.

## Licensing

This package is available under a dual-licensing model:

1. **AGPL-3.0 License** (Open Source)
   - Free for open-source projects, academic research, and personal use
   - Full terms in [LICENSE.md](LICENSE.md)

2. **Commercial License**
   - For proprietary/commercial applications
   - Allows closed-source usage
   - Details in [LICENSE_COMM.md](LICENSE_COMM.md)

For commercial licensing inquiries:
- Email: benjaminpeeters@protonmail.com
- Subject: "Commercial License - R Package WPD - [Your Use Case]"

## Development and Contributions

We welcome contributions! Please read our [Contributing Guide](CONTRIBUTING.md) for details on how to submit pull requests, report issues, and contribute to documentation.

⚠️ **Important Note**: Currently, there are known issues with legend rendering when using ggplot2 3.5.x. For optimal visualization, we recommend using ggplot2 3.4.x. You can install the recommended version using:

```r
# Install specific ggplot2 version
remotes::install_version("ggplot2", version = "3.4.3")
```

### Future Developments

We are actively working on extending the package's functionality. Here are our planned features:

* Expand database coverage:
  - Include World Inequality Database (WID)
  - Incorporate Aizenman's research data
  - Add datasets from additional researchers

* Enhance country identification flexibility:
  - Support ISO2 codes and full country names in `wp_data()`
  - Enable single-year queries in `wp_data()`
  - Add support for historical USSR ISO codes in `wp_get_category()`

* Improve user experience:
  - Develop `wp_quick()` for streamlined data visualization (e.g., balance of payments graphs)
  - Enhance `wp_info()` to display timespan and country coverage for each indicator
