---
title: "Introduction to WPD"
author: "Benjamin Peeters"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to WPD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = 'center'
)
```

# Introduction

The WPD (World Panel Data) package provides a comprehensive platform for accessing and visualizing macroeconomic data. It combines an extensive database of global economic indicators with powerful visualization tools, making it an essential resource for economists, researchers, and policy analysts.

## Key Features

- **Efficient Data Management**: Large database (~50MB) loaded on demand to optimize memory usage
- **Advanced Visualization**: Built-in plotting functions using ggplot2
- **Time Series Analysis**: Tools for handling seasonal adjustments and trend analysis
- **Flexible Data Access**: Simple interface for retrieving specific indicators and time periods

# Getting Started

## Installation

You can install the WPD package from GitHub:

```{r installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("benjaminpeeters/WPD")
```

Load the package:

```{r load}
library(WPD)
```

## Basic Usage

### Accessing Data

The main function `wp_data()` provides access to the database. Here's a simple example:

```{r basic-example}
# Get GDP data for a specific country
data <- wp_data(
  ISO = "USA",
  formula = c("GDP_C", "CU_C/GDP_C*100"),
  variable = c("GDP", "Current Account"),
  years = c(2000, 2023)
)

# Preview the data
head(data)
```

### Basic Visualization

The package includes built-in visualization functions:

```{r basic-plot}
# Create a basic time series plot
wp_plot_series(data, 
               y_axis = c("Billion USD", "% of GDP"),
               right_axis = "Current Account",
               title = "GDP and Current Account of the United States")
```

## Data Structure

The WPD database includes various categories of indicators:

1. **Economic Indicators**
   - GDP and its components
   - Trade statistics
   - Price indices
   
2. **Financial Indicators**
   - Exchange rates
   - Interest rates
   - Stock market indices

3. **Social Indicators**
   - Population statistics
   - Employment data
   - Education metrics

### Available Countries and Indicators

You can view available data using helper functions:

```{r available-data}
# View quarterly data
wp_info("Q")
```

## Advanced Features

### Seasonal Adjustment

The package provides tools for seasonal adjustment (in process).

### Custom Visualizations

Create more complex visualizations by combining multiple indicators (in process).


# Best Practices

1. **Memory Management**
   - Use date filters to limit data loading
   - Clear unused data with `rm()` when working with multiple large datasets

2. **Performance Optimization**
   - Cache frequently used datasets
   - Use built-in aggregation functions for large datasets

3. **Data Quality**
   - Check for missing values using `check_missing()`
   - Validate data consistency with `validate_data()`

# Additional Resources

- Package documentation: `help(package = "WPD")`
- GitHub repository: https://github.com/benjaminpeeters/WPD
- Bug reports: https://github.com/benjaminpeeters/WPD/issues

# Citation

If you use WPD in your research, please cite it:

```{r citation}
citation("WPD")
```

# Contributing

We welcome contributions! Please see our GitHub repository for guidelines on how to contribute to the package.

